\chapter*{Conclusions}
\lhead[\fancyplain{}{\bfseries\thepage}]{\fancyplain{}{\bfseries\rightmark}}

In this work we proposed a theoretical model for cell differentiation.
Since this biological process is governed by Gene Regulatory Networks, these networks can be modelled by Random Boolean Networks, in which each gene can be represented by  node which can be "on" or "off".
The process of differentiation is a multistable dynamical system, and involves different type of cells, but all of these start from one unique type of cell: the stem cells. The complexity of this process lays in the fact that cells "can" decide if transforming in one type of cell with respect one other, and for this reason seems pheasible that if a network for a type of cell is active, it may hinibit the network for a different type of cell.
The proposed model can give an estimate of a fitness potential for two different type of cells, according to the theory of Waddington.
This process concerns also the birth of cancer cells: cancers cell can be governed by a specific type of regulatory network, which often is inactive, but due to some external stimuli it can be activated and gives an irreversible process of production of cancer cells. This can be seen as a local minima of the Waddington potential, which is impossible to escape.
The role of noise in this model is crucial, but it is well known that biological process are indeed very dependent to external noise.

Today, constructing Gene Regulatory Networks from sequencing data is still impossible, so future studies on this model can give more details about multiple type of cells.
